Predicting Water Treatment Plant Chemical Doses

### 2025-12-10 Exploratory Analysis Ethan SchillingREAL.Rmd

```{r}
library(tidyverse)
library(knitr)
library(tidymodels)
library(dplyr)
library(Metrics)
library(kableExtra)
```

@@

for(i in 1:length(models)) {
  pred <- predict(models[[i]], newdata = test)
  performance <- rbind(performance,
                       data.frame(Model = paste0("Model ", i-1),
                                  RMSE = round(rmse(test$FeCl_dose, pred), 3),
                                  MAE  = round(mae(test$FeCl_dose, pred), 3),
                                  Description = descriptions[i],
                                  stringsAsFactors = FALSE))

@@

@@

>>>>

```{r}
library(tidyverse)
library(knitr)
library(tidymodels)
library(dplyr)
library(Metrics)
library(kableExtra)

@@

for(i in 1:length(models)) {
  pred <- predict(models[[i]], newdata = test)
  performance <- rbind(performance,
                       data.frame(Model = paste0("Model ", i-1),
                                  RMSE = round(Metrics::rmse(test$FeCl_dose, pred), 3),
                                  MAE  = round(Metrics::mae(test$FeCl_dose, pred), 3),
                                  Description = descriptions[i],
                                  stringsAsFactors = FALSE))

@@

p3 <- ggplot(test, aes(x = FeCl_dose, y = pred_model3)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Model 3: Predicted vs Actual FeCl_dose",
       x = "Actual FeCl_dose",
       y = "Predicted FeCl_dose") +
  theme_minimal()

p4 <- ggplot(test, aes(x = FeCl_dose, y = pred_model4)) +
  geom_point(color = "green") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Model 4: Predicted vs Actual FeCl_dose",
       x = "Actual FeCl_dose",
       y = "Predicted FeCl_dose") +
  theme_minimal()

p3
p4

@@

# Coagulant Dose Prediction Linear Regression

```{r}
knitr::include_graphics("Images/linear_model_table.png")
```

## Regression Models

- Created 6 models with different predictors and combinations of interaction terms.

- Dropped Turbidity as it was not a good predictor for FeCL dosing.

- Difficult to determine what factors of the dataset are driving the predictions.

### Performance

- Performed best with multiple interactions between Conductivity, Temp, and pH

- Still not great predictions many are overfit and underfit.

- Linear Structure doesn't capture the relations in our predictors

### 2025-12-10 Presentation DSCI445.Rmd

**Ethan: insert table with models attempted**

>>>>

- Created 6 models with different predictors and combinations of interaction terms.

- Used lag variables for temp, ph, conductivity, turbidity

- Dropped Turbidity as it was not a good predictor for FeCL dosing.
  - Conductivity was also dropped eventually due to being similar to Turbidity

- Difficult to determine what factors of the dataset are driving the predictions.

- RMSE and MAE seem to be all over the place dependent on which predictors are selected and the interactions.

# RMSE & MAE Table

```{r, out.width="80%", out.height="88%"}
knitr::include_graphics("Images/linear_model_table.png")
```

# Performance

- Performed best with multiple interactions between Conductivity, Temp, and pH

- Still not great predictions, many are overfit and underfit.

- Linear Structure doesn't capture the relations in our predictors, and incorrect predictions could lead to disastorus results if too much or too little chemical is used.
